Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-11-15 paper_comparison/paper_comparison/paper_comparison.Rmd
fix powre graph

What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test. The margin is largest at around $log_e(E) = 2$. An example lineup is included in Figure \ref{fig:power-poly-uniform} where none of subjects detect the actual data plot positioned at panel 10. It demonstrates that at this level of difficulty, the designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model. From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public even though it is true. At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger. At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot. Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison. The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 7, where four of 11 subjects detect it. It can be observed that a "U-shape" is presented in plot 7, but the signal is not strong enough to attract all five subjects, resulting in a visual p-value sightly above the desired significance level $\alpha = 0.05$. At $log_e(E) = 4$, the designed visual features become much clear and attractive, leading to a high percentage of rejection of the null hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible.

@@

```{r power-poly-uniform, fig.cap = "..."}
vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual p-value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual p-value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(F_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " p-value", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + geom_point(data = ~filter(.x, unique_lineup_id %in% c("poly_166", "poly_269", "poly_396", "poly_574")), aes(log_effect_size, prop_detect), alpha = 1, col = "red", size = 2) +
  theme_light(base_size = 5) +
  scale_color_brewer(palette = "Dark2", direction = -1) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") -> p

@@

```{r eval = FALSE, fig.height= 3, fig.width=4}
vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  filter(abs(log(effect_size) - 1) < 0.5) %>%
  filter(n == 100)

p

vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  filter(abs(log(effect_size) - 4) < 0.5) %>%
  filter(prop_detect < 0.25)

vi_survey %>%
  filter(unique_lineup_id == "poly_199") %>%
  View()

VI_MODEL$plot_lineup(vi_lineup$poly_159$data)

vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_abline() +
  geom_point(aes(p_value_uniform, p_value)) +
  scale_x_log10() +
  scale_y_log10()

vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  ggbeeswarm::geom_quasirandom(aes(factor(e_sigma), weighted_detect)) +
  facet_wrap(~n)
```

@@

```{r dist-power, fig.cap="..."}
vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) +
  stat_smooth(geom = "line",
              data = ~boot_y(.x, reject_dirichlet, times = 500),
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"),
              method = "glm",
              method.args = list(family = binomial),
              se = FALSE,
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value),
              aes(log_effect_size, as.numeric(value < 0.05), col = name),
              method = "glm",
              method.args = list(family = binomial),
              se = FALSE) +
  facet_wrap(~x_dist) +
  theme_light(base_size = 5)
```

```{r}
vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(b != 0) %>%
  filter(b < 128) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  facet_wrap(~x_dist) +
  ggtitle("heter")
```

>>>>

What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test. According to the example lineups included in Figure \ref{fig:power-poly-uniform}, at around $log_e(E) = 1$, the  designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model at this level of difficulty. In fact, zero out of 11 subjects detect the actual data plot positioned at panel 10. However, F-test has more than 60% chance of rejecting the null hypothesis in this case. From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public.

The margin is largest at around $log_e(E) = 2$. As shown in Figure \ref{fig:power-poly-uniform}, the visual pattern is still undetectable which is similar to $log_e(E) = 1$. The visual test power is close to zero. In contrast, F-test will reject the null hypothesis 90% of the time.

At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger. At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot. Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison. The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 20, where two of 11 subjects detect it. It can be observed that a "S-shape" is presented in plot 20, but the signal is not strong enough to attract all 11 subjects, resulting in a visual p-value sightly above $0.1$. At $log_e(E) = 4$, the designed visual features become much clear and attractive in panel 3, leading to a high percentage of rejection of the null hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible.

@@

```{r power-poly-uniform, fig.cap = "..."}
vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(F_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) + geom_point(data = ~filter(.x, unique_lineup_id %in% c("poly_166", "poly_269", "poly_396", "poly_574")), aes(log_effect_size, prop_detect), alpha = 1, col = "red", size = 2) +
  theme_light(base_size = 5) +
  scale_color_brewer(palette = "Dark2", direction = -1) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") -> p

@@

@@

```{r eval = FALSE}
vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) +
  theme_light(base_size = 5) +
  scale_color_brewer(palette = "Dark2", direction = -1) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") +
  facet_wrap(~x_dist)
```

```{r eval = FALSE}
vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  geom_point(aes(log_effect_size, prop_detect), alpha = 0.3) +
  theme_light(base_size = 5) +
  scale_color_brewer(palette = "Dark2", direction = -1) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") +
  facet_wrap(~x_dist)
```

```{r dist-power, fig.cap="..."}
vi_survey %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist)) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject_dirichlet = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)) %>%
                mutate(type = "polynomial") %>%
                mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist)) %>%
                bind_rows(mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)) %>%
                mutate(type = "heteroskedasticity") %>%
                mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete uniform", x_dist))), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  theme_light(base_size = 5) +
  scale_color_brewer(palette = "Dark2", direction = -1) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") +
  facet_grid(x_dist~type)
```

### 2022-11-15 paper_comparison/paper_comparison/paper_comparison.tex
fix powre graph

What truly impress us is the huge difference between the estimated power
of visual test and the estimated power of F-test. The margin is largest
at around \(log_e(E) = 2\). An example lineup is included in Figure
\ref{fig:power-poly-uniform} where none of subjects detect the actual
data plot positioned at panel 10. It demonstrates that at this level of
difficulty, the designed visual feature is rarely visible, making the
actual data plot indistinguishable from residual plots simulated from
the assumed model. From a communication perspective, given the fact that
the visual difference is unperceivable, the argument that non-linearity
present in the fitted model is less convincing to the public even though
it is true. At around \(log_e(E) = 3\), the margin gets smaller as the
chance of identifying the actual data plot becomes larger. At this level
of difficulty, the designed visual features are usually detectable but
it may not stand out from the lineup as other null plots may happen to
include outliers or visual patterns that are are considered to be more
attractive by human, and thus recognized as the most different plot.
Without knowing the designed visual features beforehand, it is actually
hard to identify the actual data plot by pure image comparison. The
corresponding example lineup for \(log_e(E) = 3\) shown in Figure
\ref{fig:fig:power-poly-uniform} has the actual data plot positioned at
panel 7, where four of 11 subjects detect it. It can be observed that a
``U-shape'' is presented in plot 7, but the signal is not strong enough
to attract all five subjects, resulting in a visual p-value sightly
above the desired significance level \(\alpha = 0.05\). At
\(log_e(E) = 4\), the designed visual features become much clear and
attractive, leading to a high percentage of rejection of the null
hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example
lineups of such cases. It is also interesting to observe that the power
curve of visual test behaves similarly to those inappropriate
conventional tests until the designed visual features really becomes
visible.

@@

\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1} \end{center}

>>>>

What truly impress us is the huge difference between the estimated power
of visual test and the estimated power of F-test. According to the
example lineups included in Figure \ref{fig:power-poly-uniform}, at
around \(log_e(E) = 1\), the designed visual feature is rarely visible,
making the actual data plot indistinguishable from residual plots
simulated from the assumed model at this level of difficulty. In fact,
zero out of 11 subjects detect the actual data plot positioned at panel
10. However, F-test has more than 60\% chance of rejecting the null
hypothesis in this case. From a communication perspective, given the
fact that the visual difference is unperceivable, the argument that
non-linearity present in the fitted model is less convincing to the
public.

The margin is largest at around \(log_e(E) = 2\). As shown in Figure
\ref{fig:power-poly-uniform}, the visual pattern is still undetectable
which is similar to \(log_e(E) = 1\). The visual test power is close to
zero. In contrast, F-test will reject the null hypothesis 90\% of the
time.

At around \(log_e(E) = 3\), the margin gets smaller as the chance of
identifying the actual data plot becomes larger. At this level of
difficulty, the designed visual features are usually detectable but it
may not stand out from the lineup as other null plots may happen to
include outliers or visual patterns that are are considered to be more
attractive by human, and thus recognized as the most different plot.
Without knowing the designed visual features beforehand, it is actually
hard to identify the actual data plot by pure image comparison. The
corresponding example lineup for \(log_e(E) = 3\) shown in Figure
\ref{fig:fig:power-poly-uniform} has the actual data plot positioned at
panel 20, where two of 11 subjects detect it. It can be observed that a
``S-shape'' is presented in plot 20, but the signal is not strong enough
to attract all 11 subjects, resulting in a visual p-value sightly above
\(0.1\). At \(log_e(E) = 4\), the designed visual features become much
clear and attractive in panel 3, leading to a high percentage of
rejection of the null hypothesis. Figure
\ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It
is also interesting to observe that the power curve of visual test
behaves similarly to those inappropriate conventional tests until the
designed visual features really becomes visible.

@@
